\documentclass[]{article}
\usepackage[all]{xy}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{enumitem}
\usepackage{indentfirst}
\usepackage{listings}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{tipa}
\begin{document}

\newcommand{\code}{\texttt}
\newtheorem{thm}{Theorem}
\title{Operating Systems \\ COMS W4118 \\ Lecture 11}
\author{Alexander Roth}
\date{2015 -- 02 -- 26}
\maketitle

\section{\code{select} in a Web Server}
\begin{itemize}
\item \code{select} will block until there is something to return on a socket.
\item Socket being closed on one end is an activity, so the select will return
that.
\item Context switches are something that should be avoided as they are costly
and expensive.
\item \code{select} is simpler, but slower, no concurrent execution.
\end{itemize}

\section{Domain Sockets}
\subsection{Summary of UNIX IPC so far}
\begin{itemize}
\item Shared memory, there are two kinds of shared memory, purely written memory
not backed by a file and memory that is mapped to a file.
\item File-backed mmap maps a region of memory to a file. Thus, unrelated
processes can map regions of memory to a file and access these regions.
\item Between multiple threads and processes, we have to think about
synchronization.
\item With multiple threads and processes sharing a memory region, we should use
unnamed POSIX semaphores.
\item Use a semaphore in a process context, use mutex lock, and condition
variables in a threading context.
\item Pipe returns a descriptor, while FIFO returns a name to the pipe
\item TCP sockets intended for networking, so high overhead. No boundaries as
per TCP protocol of splitting up packets.
\item UDP sockets create a pipeline, unreliable.
\end{itemize}

\subsection{UNIX Domain Sockets}
\begin{itemize}
\item Cross between pipes and sockets
\item Uses the socket API
\item Behaves like a socket
\item You can only do local communication, this is \textbf{NOT} networking.
\item Used as a socket between two processes running on the same machine.
\item Thus, we do not have any overhead related to network sockets.
\item There is a stream and datagram mode, but it will always be reliable.
\item No protocol overhead.
\item Full duplex.
\item Preserves the message boundaries unlike regular TCP.
\item Convenient, reliable, fast, and preserves message boundaries. The message
can be as long as it wants, but it will only accept the specified number of
bytes.
\end{itemize}

\end{document}
